### 2018-12-09 Revision/Response/response.tex

\clearpage

\section*{Reviewer \#1}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \referee{The authors find that the uncertainty in the ignition delay time is governed
by the uncertainty in H+O~2$<$=$>$~O+OH.
The connection between this reaction and the ignition delay time is so overwhelming that it
is probably responsible for the 1-dimensional active subspace reduction. Once the sensitivity
indices are computed, the one associated with this key reaction is largest.
To a combustion scientist, the fact that ignition delay time (and the uncertainty in the 
ignition delay time when pre-exponentials or activation energies are changed) is governed 
by H+O2~$<$=$>$~O+OH is obvious. This is THE key reaction in combustion kinetics with 10s 
(if not 100s) of studies dedicated to it (from the beginning of the field…). This is textbook k knowledge.
E.g. Figure 4 in Hong et al. (PCI 33, 2011, p. 309-316) shows that, indeed, this is the key 
reaction in sensitivity (by a large margin…). That figure is a "sensitivity analysis" 
which one can accomplish easily.
So, if this is the case, is the authors' reduction framework working because the nonlinear system
is really "simple" in its response to varying parameters}

The complexity of the H$_2$/O$_2$ reaction kinetics application in the revised manuscript has been significantly
enhanced by increasing the degree of uncertainty in the rate parameters, and considering additional 
uncertainty in the
initial pressure, temperature, and stoichiometry of the system. As a result, the non-linearity in the system
response is observed to increase substantially as seen in Figures 3 and 6 unlike a straight line as 
observed earlier. Our sensitivity analysis results in Figure 4 using the proposed framework based on the
active subspace methodology are shown to be consistent with our earlier results using the derivative-based
global sensitivity analysis (Ref. 35{\color{blue} Update}). The sensitivity results for the higher-dimensional problem in the
revised manuscript reveal that the ignition delay is largely impacted by
the uncertainty in the pre-exponents associated with reactions 1, 9, and the initial temperature. Additionally,
we observe significant sensitivity towards the rate parameters associated with reaction 15 and the pre-exponent
associated with the rate of reaction 17. 
Therefore, depending upon the considered prior uncertainty in the individual rate parameters for the 19
reactions, and the range of values of the initial conditions, the proposed framework helps identify the
key contributors to the uncertainty in the ignition delay. The analysis accounts for individual 
contributions of the parameters as well as contributions due to their interactions with other parameters to the 
output uncertainty in the system response. Through this relatively more complex
problem, we have demonstrated the robustness of the proposed framework.

@@

As mentioned in Section 6
(Summary and Discussion), the proposed framework is agnostic to the choice of the system and could be extended
to other kinetics applications as long as the system response is continuously differentiable in the domain of
the inputs and is therefore applicable to a wide variety of applications.

\item \referee{Related to 1) above, it is of paramount importance that the authors consider, at least, 
methane, for which I expect less of a spectral gap in the matrix C. Ideally, more complex fuels with 
more nuanced responses.
In other words, if the method works well for a simple case, what is its value to the community?
The authors must try this out on a much more complex hydrocarbon/oxygen system.}

A more complex problem as suggested by the reviewer would manifest itself into an increased number of
uncertain parameters and/or an increased amount of non-linearity in the system response. Therefore, while
we did not implement the framework to a new system, we have significantly increased the non-linearity of the
response by increasing the degree of uncertainty in the rate parameters. In addition to the rate parameters,
we perturb the initial conditions to investigate their impact on the response individually as well as due to
their interaction with the rate parameters. Hence, we have also increased the dimensionality of the problem
for the higher dimensional case from 33 to 36. More importantly, the impact of initial conditions on the
variability in the system response could provide useful insights to combustion scientists interested in 
reaction kinetics for a range of applications.

@@

As expected by the reviewer, the resulting spectral gap between the first two eigenvalues is observed to
decrease for this relative more complex scenario as shown in Figure~1 in the revised manuscript. 
The proposed framework has been used effectively to compute the active subspace, construct a low-dimensional
surrogate, and perform a global sensitivity analysis in this case.

\item \referee{The authors report ignition delay times of 0.1 s (?). See for example Fig. 7 and Tab. 2.
Are these seconds (s)? If this is the case, what kind of conditions (temperature/pressure/stoichiometry)
are the authors considering? They seem very unphysical if the yield an ignition delay time of O(0.1 s).
I could not find any clarification of the initial conditions for the ignition calculations.
This is again a symptom of the disconnect between the authors' aims (a novel method, potentially very useful to the combustion community) and the combustion community's aims (more understanding and tools that can be of help in combustion kinetics).}

Depending upon the initial conditions, the ignition delay can range from $\mathcal{O}(10^{-3})$
to $\mathcal{O}(10^3)$ seconds for hydrogen-air mixtures as shown by Starostin et al. in~\cite{Starostin:2011}.
In this work, the initial pressure ($P_0$), temperature ($T_0$), and
the equivalence ratio~($\Phi_0$) considered in the computations are 1~atm, 900~K, and 2.0
respectively. These values have been reported in the revised manuscript in the beginning of
Section 5 on page 15.
In this revision, we perturb the initial conditions to study their impact
on ignition delay estimates. The mean estimate of the ignition delay using data at 10$^4$ random samples in
the 36-dimensional space is found to be approximately 0.15 as reported in Table 2 on page 18 in the
revised manuscript. As mentioned in the paper, we have used 
the TChem software for model predictions and based on our recent communication with the developers at
Sandia National Laboratories, these estimates seem reasonable.

@@

\item \referee{p.3, line 36. I would call this the ``temperature exponent".}

The reviewer's suggestion has been incorporated in the revised manuscript on Page 3.

@@

Thanks for this comment. As the referee notes, while the spectral 
truncations in active subspace methodology resemble aspects of CSP analysis,
the goals and approaches of these methods are substantially different. However, 
we are aware that CSP analysis is well-known in combustion 
community, and therefore, have included a brief mention 
of CSP in the introduction, clarifying the differences between 
the active subspace method and CSP.

\item \referee{p. 16, Fig. 6. y-axis label. Probably $\lambda_i/\lambda_0$ as goes to 1 for i=1? Typo?}

It is in fact $\log(\lambda_i/\lambda_0)$. The typo has been corrected.

%Minor comments by the reviewer have also been addressed in the revised manuscript.

@@

\end{enumerate}

\section*{Reviewer \#2}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \referee{There is a lack of quantitative discussion of the expense of
regression-vs-perturbation, in terms of the actual number of model simulations
required. The work seems to suggest that perturbation works better than
regression, but then it is a direct and clear consequence of the fact that
local approximation is supposed to be better than a global one. Ideally, much
more quantitative comparison needs to be made here before reaching any
meaningful conclusions.}

As suggested by the reviewer, we have focused on a quantitative
comparison in terms of the number of samples and associated model runs required
to obtain a converged active subspace for the two approaches. In addition to
convergence characteristics, we have also compared the two approaches for their
predictive accuracy for the same computational effort for the 19-dimensional
problem as well as the 36-dimensional problem. In the revised work, we have
increased the dimensionality of the higher-dimensional case from 33 to 36.
Specifically, we have enhanced our analysis to study the impact of
perturbations in the initial conditions: pressure, temperature, and
stoichiometry on the system response i.e. ignition delay.

@@

\item \referee{Another fundamental flaw is related to the additive treatment of
the uncertainty of the pre-exponential factor A. In most works dealing with
uncertainty in chemical kinetics, the uncertainty form is multiplicative or,
equivalently, log(A) is the underlying fundamental parameter with additive
uncertainty range. In the current form of the uncertainty, the problem is too
benign, and barely carries any practical usability for the practitioners.}

The reviewer has made a valuable suggestion. Therefore, in the
revised manuscript, we have updated the computations for the active subspace
using perturbation and regression by considering the uncertainty in $\log(A_i)$
as opposed to $A_i$.  As a result, the non-linearity of the system response is
observed to increase significantly. As shown in Figures~3 and 6, the
variability of the ignition delay in the active subspace is no longer a
straight line as observed earlier. Hence, we have substantially increased the
complexity of the problem by increasing the non-linearity in the system
response as well as its dimensionality to illustrate the robustness and the
applicability of the proposed framework for a wide range of kinetics
applications.

\item \referee{Why not use the word 'active' in the title, since active subspace is essentially
what has been done here?}

The title has been updated as ``Active subspace-based dimension
reduction for chemical kinetics applications with epistemic uncertainty".

@@

\item \referee{3.36 $n_i$ is index or power? Is this an established term for
the power of T?  Also, what are the values of $n_i$'s?}

In the revised manuscript, we denote $n_i$ as the `temperature
exponent' as suggested by the other reviewer. Yes, it is an established
quantity and its value in addition to the rate constants for the 19 reactions
in the H$_2$/O$_2$ mechanism is taken from [Yetter et. al., 1991] and
implemented in the TChem software used for model predictions in this work. We
have clarified this on Page 4 in the revised manuscript.

@@

\item \referee{3.51 'rate of temperature' or 'temperature'?}

The ignition delay is considered as the time required for the {\it rate of temperature}
increase to exceed a threshold, in this work.

@@

\item \referee{4.57 not clear where the number 33 comes from since we have 19
reactions. I see it is explained at the beginning of Section 5, but it should
be noted here briefly, too.}

We mention, in Page 4 (just below Table 1) in the introduction, that only the
activation energies with a non-zero nominal value are considered as uncertain,
in the higher-dimensional case.  Since the uncertainty in the activation energy
is considered only in the higher-dimensional case presented in Section 5, we
have specified the reactions for which it is 0 in the beginning of that
section.

@@

\item \referee{6.19 distribution-$>$density function}

The revised manuscript has been updated to include the reviewer's suggestion.

@@

\item \referee{6.38: typo, $w_r$ should be $w_1$}

The typo has been fixed in the revised manuscript.

\item \referee{7.23-27: why 'statistical independence' and not just 'independence'? Is there any other 
type of independence?}

We agree; we have removed 'statistical' from that part.

\item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is
just the negative log-PDF, I am not sure if it is necessary to bring up a
Boltzmann distribution language here.  Also, it could be confused with
Maxwell-Boltzmann distribution, which is a very specific PDF form.}

Thanks for
this comment. We followed the theory in ref [35] in the manuscript, where the
authors derive Poincare type inequalities for functions of iid random inputs
distributed according to Boltzmann probability laws.  We point out that the
uniform distribution does not belong to this class.  The required Poincare type
inequality for the case of uniform distribution was derived in an earlier paper
of Sobol' and Kucherenko (ref [34]).  That was the reason for assuming the
distribution laws of the random inputs are either \emph{uniform} or belong to
the Boltzmann class, in section 3. An important class of Boltzmann measures are
the so called log-concave distributions such as Normal, Exponential, Beta, and
Gamma distributions. These points have been clarified in the revised
manuscript.

@@

Additionally, we share the reviewer's concern about the potential confusions
some readers might face when seeing the mention of Boltzmann. We have removed
the mention of Boltzmann measures from the introduction, and now only refer to
Boltzman measures in Section 3, where the theoretical results are discussed.

\item \referee{7.40: Explain what E and V are with respect to in Eq (8).}

That equation as well as the following description of the conditional 
expectation and the variance has been updated to address the reviewer's comment.

@@

\item \referee{7.50: Just say in words that it is the i-th diagonal element of matrix C.}

Following the reviewer's suggestion, we have updated the text in the revised manuscript to
clearly note that the DGSM $\nu_i(f)$ is the $i^{\text{th}}$ diagonal element in the matrix, $\mathbf{C}$
on Page 8.

@@

\item \referee{10.56: what if there is no eigenvalue gap of O(100)?}

A large gap in the eigenvalue spectrum typically indicates that a low-dimensional (1--2) active subspace 
would sufficiently capture the variability in the model output due to uncertainty in the inputs. However,
in the revised manuscript, we have shown that the variability in the ignition delay is reasonably
approximated with a 1-dimensional active subspace even though the gap in the first two eigenvalues
is found to be O(10). Hence, a smaller gap in the eigenvalue spectrum might still lead to a reasonable
accuracy with a low-dimensional active subspace depending upon the relationship between the
system response and the inputs. However, it is also possible that we might need to a consider a higher 
dimensional active subspace to be able to approximate the variability in the response with reasonable
accuracy in situations where the gap is small.

@@

\item \referee{11.33: Alg. 1: should have $\beta$ as input, too?}

The reviewer is correct. Algorithm 1 in the revised manuscript has been updated to include $\beta$
as an input.

\item \referee{11.42: Which norm is used? Any intuition why squares of W are used, and not just W?}

We have used the L$^2$ norm to estimate the relative change in the squared values of corresponding
eigenvector components between successive iterations. The algorithm in the revised manuscript has been updated 
to specify the norm. 
We consider square of individual components due to the fact the both W and -W contain legitimate eigenvectors 
of C. Alternatively, we could have also used the modulus value of the components to assess convergence.

@@

\item \referee{11.45: what is max over?}

The max is over all eigenvectors in the active subspace denoted by $j$. This has been specified in
the updated algorithm in the revised manuscript.

@@

\item \referee{12.29: 'higher-order' should be replaced with higher-index or smaller-magnitude, to avoid confusion.}

The reviewer's suggestion has been incorporated in the revised manuscript.

@@

\item \referee{Figs 2a, 3 and 4 demonstrate the regression-based gradient computation is working
similarly to the perturbation based one. However, the comparison is really qualitative and does
 not provide much insight. Clearly, regression will approximate the gradient less accurately then
a local perturbation, but the latter has a price associated with it as new simulations are needed
(a factor of d+1). It would have been great to see how accurate each method is for comparable
expense, or vice-verse, how expensive each approach is for a comparable accuracy. I know there
is a statement about this at the end of Section 4, but it is neither sufficient nor quantitative.
Same goes for statements in Section 5.2 - it was not clear whether regression- and
perturbation- based approaches had comparable expenses (in terms of model evaluations) or not.}

Based on the reviewer's feedback, in the revised manuscript, the analysis has been
updated to
compare the perturbation and regression approaches for their convergence and predictive 
accuracy using the same amount of computational effort. As a result, the comparative 
analysis is much more conclusive with regards to the performance of the two approaches.

\item \referee{For Figure 8b, the reasoning why E$_{15}$ was not captured by the regression approach,
is not satisfactory. I think the authors could have gone a bit deeper in order to explain this
glaring miss of the regression-based approach. There is some discussion in Section 6, but it is
equally unsatisfactory.}

Our latest analysis in the revised manuscript has revealed that the two approaches
in fact yield consistent results
for sensitivity analysis even for the 36-dimensional case with increased non-linearity
when using the same amount of
computational effort or the same number of model evaluations. Earlier, the two approaches were
compared for the same number of samples, N. However, the number of model runs in the case of
perturbation was N(d+1); d is the number of uncertain parameters and inputs to the model. Whereas,
the number of model runs in the case of regression was only N. Therefore, it can be said that
the accuracy of the regression-based approximation of the gradient can be increased to the 
extent of perturbation-based approach by increasing the sample size.

@@

\item \referee{14.4-8: the near-linearity simply makes the problem extremely benign, nearly any
reasonable method would have captured such behavior. It would have been great to consider more
complicated case (more than simply going from 19 to 33 parameters), perhaps looking at logA 
perturbation instead of A would help? Also consider more than 1$\%$ 
perturbations to the input. Otherwise, the problem is too simple and does not do justice to the method.}

We have made a sincere effort to incorporate these valuable suggestions by the 
reviewer in the revised manuscript. 
Specifically, we perturb $\log(A_i)$ instead of $A_i$ 
which is observed to significantly increase
the non-linearity of the system response. In addition to the rate parameters, we analyze the impact
of perturbing the initial conditions. Moreover, as suggested by the reviewer, we have
increased the magnitude of the perturbations from 1$\%$ to 2$\%$ to illustrate the applicability
and robustness of the proposed approach to more complex scenarios.

@@

\item \referee{14.8: here and in a few other places, I have an issue with statements like
active subspace is 1-dimensional'. To me, active subspace is as many dimensional as one likes,
so the statement should be changed to something like '1-dimensional active subspace captures
most of the variance/uncertainty/dynamics etc..', or '1-dimensional active subspace provides
very accurate approximation...'.}

The reviewer's suggestion has been incorporated in the revised manuscript.

\item \referee{14.57: even if nominal is zero, one could perturb them, right? 
Or it is actually indicating the fact that there should be no exponential factor whatsoever.}

The latter statement is correct. We discussed this in our group extensively,
and also consulted with the developers of the TChem software at Sandia National
Laboratories, and decided not to perturb the activation energies with a 0
nominal value since the exponential factor is not considered for these
reactions, as mentioned by the reviewer.

@@

\item \referee{18.35: the remark in the parentheses did not make sense to me.}

Our goal was to list some efficient alternatives to finite-difference gradient
computations. We have modified that part to make things clearer.

### 2018-12-09 Revision/Paper/assa.tex
...

\section{GSA measures and their links with active subspaces}
\label{sub:gsa}
Consider a function $f = f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
While the active subspace framework described above does not make any assumptions
about statistical independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
the classical 
framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} 
assumes that the inputs
are statistically independent. While extensions to the cases 
of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007},
 we limit the discussion in this section to the
case of random inputs that are statistically independent and are 
either uniformly distributed or
distributed according to the Boltzmann probability distribution.
Note that a measure $\mu$
on $\R$ is referred to as a Boltzmann measure if it is 
absolutely continuous with respect to the Lebesgue measure  
and admits a density  of the form $\pi(x) = C \exp\{-V(x)\}$,
where $V$ is a continuous function and $C$ a normalization 
constant~\cite{Lamboni:2013}. \rebut{
An important class of Boltzmann distributions are the so called log-concave
distributions, which include Normal, Exponential, Beta, Gamma, Gumbel, and
Weibull distributions. Note also that the uniform distribution does not fall under
the class of Boltzmanm distributions.~\cite{Lamboni:2013}.}

>>>>

\section{GSA measures and their links with active subspaces}
\label{sub:gsa}
Consider a function $f = f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
While the active subspace framework described above does not make any assumptions
about independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
the classical 
framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} 
assumes that the inputs
are statistically independent. While extensions to the cases 
of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007},
 we limit the discussion in this section to the
case of random inputs that are statistically independent and are 
either uniformly distributed or
distributed according to the Boltzmann probability distribution.
Note that a measure $\mu$
on $\R$ is referred to as a Boltzmann measure if it is 
absolutely continuous with respect to the Lebesgue measure  
and admits a density  of the form $\pi(x) = C \exp\{-V(x)\}$,
where $V$ is a continuous function and $C$ a normalization 
constant~\cite{Lamboni:2013}. \rebut{
An important class of Boltzmann distributions are the so called log-concave
distributions, which include Normal, Exponential, Beta, Gamma, Gumbel, and
Weibull distributions. Note also that the uniform distribution does not fall under
the class of Boltzmanm distributions.~\cite{Lamboni:2013}.}

### 2018-12-09 Revision/Paper/conc.tex
..
...

Active subspace computation requires repeated evaluations of the gradient of
the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
namely, perturbation-based (e.g. finite difference - used in this work, adjoint-based
methods, etc.) and regression-based. Both approaches were shown to yield consistent
results for the 19-dimensional
problem wherein only the pre-exponents were considered to be uncertain.
\rebut{It was observed that the computational effort required to obtain a 
converged active subspace was comparable for the two approaches. However,
the predictive accuracy of the perturbation approach was observed to be
slightly higher. Moreover, a 1-dimensional active subspace was shown to
reasonably approximate the uncertainty in the ignition delay.}
Additionally, the activity scores were also shown to be consistent with the
screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
iterative procedure was adopted to enhance the computational efficiency.

@@

Therefore, in general, perturbation-based approaches should be preferred for 
active subspace computation especially in situations where efficient gradient
computation techniques (adjoints-based, automatic differentiation, etc.) are
feasible. The regression-based approach can be explored in situations involving intensive
simulations where the perturbation-based methods are intractable.
%
%and the goal is obtaining rough estimates of the statistics of the QoI as
%opposed to a more detailed analysis such as those involving parametric
%sensitivity.  
%
We also mention that alternate regression based approaches such as ones based
on computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{Constantine:2017a}.  The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

>>>>

Active subspace computation requires repeated evaluations of the gradient of
the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
namely, perturbation and regression. Both approaches were shown to yield
consistent results for the 19-dimensional problem wherein only the
pre-exponents were considered to be uncertain.  \rebut{It was observed that the
computational effort required to obtain a converged active subspace was
comparable for the two approaches. However, the predictive accuracy of the
perturbation approach was found to be relatively higher. Moreover, a
1-dimensional active subspace was shown to reasonably approximate the
uncertainty in the ignition delay.} Additionally, the activity scores were also
shown to be consistent with the screening metric estimates based on DGSMs
in~\cite{Vohra:2018}. An iterative procedure was adopted to enhance the
computational efficiency.

@@

Based on our findings, the perturbation approach is preferable for active
subspace computation; moreover, this can be made extremely efficient, when more
efficient gradient computation techniques (adjoints-based, automatic
differentiation, etc.) are feasible. The regression-based approach can be
explored in situations involving intensive simulations where gradient
computation is very challenging.

%
%and the goal is obtaining rough estimates of the statistics of the QoI as
%opposed to a more detailed analysis such as those involving parametric
%sensitivity.  
%
We also mention that alternate regression based approaches such as ones based
on computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{Constantine:2017a}.  The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

### 2018-12-09 Revision/Paper/intro.tex
...

We seek to understand the impact of uncertainty in the
rate-controlling parameters, pre-exponents~($A_i$'s) 
and the activation energies~($E_{a,i}$'s) 
\rebut{as well as the initial pressure, temperature, and the
equivalence ratio} on the ignition delay. The \rebut{$\log(A_i)$'s} associated with all
reactions and the $E_{a,i}$'s with non-zero nominal estimates
are considered to be uniformly distributed about their nominal estimates provided
in~\cite{Yetter:1991}. \rebut{Temperature exponent, $n_i$ for each reaction is fixed to
its nominal value, also provided in~\cite{Yetter:1991}.
The initial conditions are also considered to be uniformly
distributed about their respective aforementioned values.} 
The total number of uncertain inputs is \rebut{36} which makes
the present problem computationally challenging due to the large number of 
uncertain parameters \rebut{in addition to the initial conditions}.  
To address this challenge, we focus on reducing the dimensionality
of the problem by computing the active subspace.
This involves repeated evaluations of the gradient of a model output with
respect to the input parameters. Several numerical techniques are available
for computing the gradient. Perturbation-based techniques
include finite differences and more advanced methods involving
adjoints~\cite{Jameson:1988,Borzi:2011,Alexanderian:2017}. The
adjoint-based method requires a solution of the state equation (forward solve)
and the corresponding adjoint equation. Hence, it is 
limited by the availability of an adjoint solver. Additional model evaluations
at neighboring points are required if finite difference is used which increases
the computational effort. Regression-based techniques on the other hand,
aim to estimate the
gradient by approximating the model output using a regression fit. These
are computationally less intensive than the former. However, as expected,
there is a trade-off between computational effort and accuracy in the two
approaches for estimating the gradient.

In this work, we adopt an iterative strategy to reduce the computational effort
associated with active subspace computation. Moreover, we explore two approaches
for estimating the gradient of the ignition delay with respect to the uncertain
rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism \rebut{as well as
the initial conditions: $P_0$, $T_0$, and $\Phi_0$. Note that the equivalence ratio
corresponding to the initial molar ratios of $H_2$ and O$_2$ is denoted as $\Phi_0$.}
As mentioned above, the
first approach uses finite difference to estimate the gradient and is regarded as
the perturbation approach. On the other hand, the second approach 
follows from~\cite{Constantine:2015} and involves repeated
regression-fits to a subset of available model evaluations, and is 
regarded as the regression approach in this work.

>>>>

We seek to understand the impact of uncertainty in the
rate-controlling parameters, pre-exponents~($A_i$'s) 
and the activation energies~($E_{a,i}$'s) 
\rebut{as well as the initial pressure, temperature, and the
equivalence ratio} on the ignition delay. The \rebut{$\log(A_i)$'s} associated with all
reactions and the $E_{a,i}$'s with non-zero nominal estimates
are considered to be uniformly distributed about their nominal estimates provided
in~\cite{Yetter:1991}. \rebut{Temperature exponent, $n_i$ for each reaction is fixed to
its nominal value, also provided in~\cite{Yetter:1991}.
The initial conditions are also considered to be uniformly
distributed about their respective aforementioned values.} 
The total number of uncertain inputs is \rebut{36} which makes
the present problem computationally challenging due to the large number of 
uncertain parameters \rebut{in addition to the initial conditions}.  
To address this challenge, we focus on reducing the dimensionality
of the problem by computing the active subspace.
This involves repeated evaluations of the gradient of a model output with
respect to the input parameters. Several numerical techniques are available
for computing the gradient, such as 
finite differences and more advanced methods involving
adjoints~\cite{Jameson:1988,Borzi:2011,Alexanderian:2017}. The
adjoint-based method requires a solution of the state equation (forward solve)
and the corresponding adjoint equation. Hence, it is 
limited by the availability of an adjoint solver. Additional model evaluations
at neighboring points are required if finite difference is used which increases
the computational effort. 
Regression-based techniques, which can be suitable for active subspace computations,
on the other hand, aim to estimate the gradient by approximating the model
output using a regression fit.  
These are computationally less intensive than
the former. However, as expected, there is a trade-off between computational
effort and accuracy in the two approaches for estimating the gradient.

In this work, we adopt an iterative strategy to reduce the computational effort
associated with active subspace computation. Moreover, we explore two
approaches for estimating the gradient of the ignition delay with respect to
the uncertain rate-controlling parameters: pre-exponent, $A_i$ and the
activation energy, $E_{a,i}$, for $i^{\text{th}}$ reaction in the H$_2$/O$_2$
mechanism \rebut{as well as the initial conditions: $P_0$, $T_0$, and $\Phi_0$.
Note that the equivalence ratio corresponding to the initial molar ratios of
$H_2$ and O$_2$ is denoted as $\Phi_0$.} The first approach uses finite
difference to estimate the gradient and will be referred to as the
\emph{perturbation} approach throughout the article.  The second approach
is adapted from~\cite[Algorithm 1.2]{Constantine:2015} and involves
repeated regression-fits to a subset of available model evaluations, and is
regarded as the regression approach in this work.
